The semantic drift of quotations in blogspace: a case study in short-term cultural evolution

### 2016-06-30 brains-introduction.tex
feat/methods: rewrite up to 'word-level measures' (excluded)
Also fix some typos and formatting things in the introduction (using
\hbox{} for "i.e.", "e.g.", "vs.").

Since the very beginnings of both social science and psychology, scientists have tried to capture the way cognition and culture influence each other.
%the reciprocal influences of cognition and culture in meaningful ways.
While this has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by e.g. Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980}), today's discussion is mostly structured by proponents from cognitive science.

@@

The theory of epidemiology of representations proposes a unifying framework for all these works by recasting them as questions of spread and transformation of representations:
these are alternatively located in the mind ("mental representations" in Sperber's terminology), or in the outer world ("public representations") as expressions of mental representations in diverse cultural artifacts (pieces of text, utterances, pictures, building techniques, etc.).
A human society is then modeled as a large dynamical system of people constantly interpreting public representations into mental representations, and producing new public representations based on what they have previously interpreted.
Two key points are that (a) transmission is not reliable (representations change significantly each time they are interpreted and produced anew, as opposed to e.g. memetics), and (b) the reciprocal influences of cognition and culture can be captured by studying the evolution of public representations themselves, which is what the studies cited above are doing.

@@

To show this we analyze the way quotes in blogs and media outlets are modified when they are copied from website to website.
These public representations should normally not change as they spread on the Web (as opposed to more elaborate expressions or opinions, not identified as quoted utterances), but empirical observation shows that they are in fact occasionally transformed~\citep{simmons_memes_2011}:
authors spontaneously transform quotes, not only cropping them but also replacing words, when in fact they are implicitly required to copy them exactly.
We can therefore assume that most transformations, especially the simple ones, are the result of automatic (i.e. hard to control) low-level cognitive biases of the authors.

@@

The study of \newtext{cultural evolution on the part of cognitive science} emerged only recently.
\newtext{While formal models of cultural transmission appeared with the development of dual inheritance theory~\citep{cavalli-sforza_cultural_1981,boyd_culture_1985} and have included the notion of cultural attractor since then~\citep{claidiere_role_2007,claidiere_how_2014}, collecting data to test and iterate over such models has been more challenging.
The first method mentioned above consists in rebuilding the history of a given type of representation by compiling anthropological or historical works on the subject (as for instance \citealp{morin_how_2013}, and \citealp{miton_universal_2015}, have done).\footnote{
Critics like \citet{ingold_trouble_2007}, however, have noted that such quantitative use of works from the social sciences approach risks overlooking the ontological debates in history and anthropology over the way to interpret such data.
}
A second approach uses cultural evolution experiments in the laboratory, with an array of methods reviewed by \citet{mesoudi_multiple_2008}.
Transmission chains, in particular, have been used extensively to study the evolution of human language~\citep[see][for a review]{tamariz_cultural_2016}.
Other recent examples include studies of the evolution of simple audio loops through consumer preference~\citep{maccallum_evolution_2012}, the emergence of structure in visual patterns transmitted by baboons~\citep{claidiere_cultural_2014}, and the amplification of risk perception through chains of casual conversation~\citep{moussaid_amplification_2015}.}

\newtext{Research on online content points to a third approach to this question.}
By investigating the transformations of quotations in a large corpus of US blog posts and online news stories initially collected and studied by \citet{leskovec_meme-tracking_2009}, \citet{simmons_memes_2011} and later \citet{omodei_multi-level_2012} show that even for quotations, a type of public representation that should change the least when transmitted on the Web, it is still possible to witness significant transformations.
These studies focus on the influence of the quotation source (e.g. news outlet vs. blog) or of the surrounding public space (e.g. quotation frequency in the corpus), and suggest diffusion-transformation models to capture the dynamics of the population of quotations.
But the cognitive features which may determine or, at least, influence these transformations, are overlooked.
\newtext{Yet cognitive and linguistic features have been used in studies not involving transformation: \citet{danescu-niculescu-mizil_you_2012}, for instance, show that particularly memorable quotations (taken from movie scripts in this case) use more distinctive words and have more common syntax than less memorable quotations; they are also the quotes that adapt best to new contexts of use.
One source of ideas to study the transformations of such quotes, then, might be the psycholinguistic literature studying word and sentence recall.}

@@

Recall of word lists provides a situation that is easier to explore fully compared to recall of complete sentences, and has been extensively studied.
In particular, the Deese, Roediger, and McDermott paradigm (introduced by \citealp{deese_prediction_1959}, and later popularized by \citealp{roediger_creating_1995}) has shown that it is possible to construct lists of words which reliably create the false memory of an external word related to those in the list.
This is done by using lists of words produced by free association from the target intrusion word;
the intruding recall then happens with probability nearly proportional to the average strength of semantic association between the intruding word and the words in the list.
A sizable literature studies this type of task with varying complexities in the design of the lists, a good review of which is given by \citet{zaromb_temporal_2006}.
One notable effect is that the semantic relations between words greatly influence, and correlate to, the order in which subjects recall a list~\citep{tulving_subjective_1962,howard_when_2002}, and that this reordering of items improves subjects' repeated recalls~\citep{tulving_subjective_1966}
The frequency and type of intrusions in lists of random words are also influenced by associations created by the presentation of previous lists~\citep{zaromb_temporal_2006};
indeed the question of how such temporal associations (contributing to contextual information retrieval in recall) interact with the prior semantic associations (contributing to associative information retrieval) of subjects is at the center of many of these studies.

@@

Given this complexity we decided to focus on more aggregate effects, where variations of the conditions in which sentences are read and produced have a chance of being statistically smoothed out.\footnote{Aside from our lack of control on the precise conditions of encoding and recall in our data set, the analysis techniques mentioned above are better suited when the data consists of a high number of measures over a smaller number of lists (in which case it makes sense to ask e.g. what proportion of intrusions come from prior lists).
As is explained further down however, our data set is shaped the opposite way: a great number of sentences, with only very few to no measures at all on each sentence.
}
If a cognitive bias in the substitution of words manifests itself with simple measures, then it will be worth applying predictive models of the substitution process in further research.

Lexical features, then, are obvious well-studied word measures that can be analyzed in aggregate.
Indeed word frequency~\citep[see][for a review]{yonelinas_nature_2002}, age-of-acquisition~\citep{zevin_age_2002}, number of phonemes~\citep[see for instance][]{rey_phoneme_1998,nickels_dissociating_2004}, and phonological neighborhood density~\citep{garlock_age--acquisition_2001} to name a few, all have known effects on word recognition or production.
More complex features based on word networks built from free association or phonological data have also been analyzed:
\citet{nelson_how_2013} for instance, show the importance of clustering coefficient in such a network by studying the role it plays in a variety of recall and recognition tasks (extralist and intralist cuing, single item recognition, and primed free association).
\citet{chan_network_2010} show that pictures are named faster and with fewer mistakes when they have a lower clustering coefficient in an underlying phonological network.
\citet{griffiths_google_2007} analyze a task where subjects are asked to name the first word which comes to their mind when they are presented with a random letter from the alphabet.
The authors show that there is a link between the ease of recall of words and their authority position (pagerank) in a language-wide semantic network built from external word association data~(\citealp{austerweil_human_2012}, further develop this tool to give a parsimonious account of the fact that related words are often retrieved together from memory).

@@

On the whole, research on lexical features hints towards two antagonistic types of effects.
On one hand, part of the literature shows that recall is easier for the least "awkward" words;
those whose age of acquisition is earlier, length is smaller, semantic network position is more central --- this is particularly true in tasks where participants are asked to form spontaneous associations or utter a word in response to a given signal.
On the other hand, when the task consists in recognizing a specific item in a list, "awkward" words are actually more easily remembered, possibly as they are more informative and plausibly more discernible~\citep[see again][for a review]{yonelinas_nature_2002}.
The jury is still out as to whether reformulation alteration, that is spontaneous replacement of words when asked to repeat a given utterance, is rather of the former or latter sort.
We also aim to shed some light on this debate, considering oddness as a dimension of the purported fitness of utterances.

>>>>

Since the very beginnings of both social science and psychology, scientists have tried to capture the way cognition and culture influence each other.
%the reciprocal influences of cognition and culture in meaningful ways.
While this has been the subject of intense debate in the social sciences in the 20th century (starting with Durkheim's initial works, \citeyear{durkheim_les_1912}, later tackled in earnest by \hbox{e.g.} Mauss' \emph{Techniques of the Body}, \citeyear{mauss_les_1936}, Giddens' \emph{Structuration Theory}, \citeyear{giddens_constitution_1984}, and Bourdieu's \emph{Sens Pratique}, \citeyear{bourdieu_sens_1980}), today's discussion is mostly structured by proponents from cognitive science.

@@

The theory of epidemiology of representations proposes a unifying framework for all these works by recasting them as questions of spread and transformation of representations:
these are alternatively located in the mind ("mental representations" in Sperber's terminology), or in the outer world ("public representations") as expressions of mental representations in diverse cultural artifacts (pieces of text, utterances, pictures, building techniques, etc.).
A human society is then modeled as a large dynamical system of people constantly interpreting public representations into mental representations, and producing new public representations based on what they have previously interpreted.
Two key points are that (a) transmission is not reliable (representations change significantly each time they are interpreted and produced anew, as opposed to \hbox{e.g.} memetics), and (b) the reciprocal influences of cognition and culture can be captured by studying the evolution of public representations themselves, which is what the studies cited above are doing.

@@

To show this we analyze the way quotes in blogs and media outlets are modified when they are copied from website to website.
These public representations should normally not change as they spread on the Web (as opposed to more elaborate expressions or opinions, not identified as quoted utterances), but empirical observation shows that they are in fact occasionally transformed~\citep{simmons_memes_2011}:
authors spontaneously transform quotes, not only cropping them but also replacing words, when in fact they are implicitly required to copy them exactly.
We can therefore assume that most transformations, especially the simple ones, are the result of automatic (\hbox{i.e.} hard to control) low-level cognitive biases of the authors.

@@

The study of \newtext{cultural evolution on the part of cognitive science} emerged only recently.
\newtext{While formal models of cultural transmission appeared with the development of dual inheritance theory~\citep{cavalli-sforza_cultural_1981,boyd_culture_1985} and have included the notion of cultural attractor since then~\citep{claidiere_role_2007,claidiere_how_2014}, collecting data to test and iterate over such models has been more challenging.
The first method mentioned above consists in rebuilding the history of a given type of representation by compiling anthropological or historical works on the subject (as for instance \citealp{morin_how_2013}, and \citealp{miton_universal_2015}, have done).\footnote{
Critics like \citet{ingold_trouble_2007}, however, have noted that the quantitative use of data imported from the social sciences risks overlooking the ontological debates in history and anthropology over the way to interpret such data.
}
A second approach uses cultural evolution experiments in the laboratory, with an array of methods reviewed by \citet{mesoudi_multiple_2008}.
Transmission chains, in particular, have been used extensively to study the evolution of human language~\citep[see][for a review]{tamariz_cultural_2016}.
Other recent examples include studies of the evolution of simple audio loops through consumer preference~\citep{maccallum_evolution_2012}, the emergence of structure in visual patterns transmitted by baboons~\citep{claidiere_cultural_2014}, and the amplification of risk perception through chains of casual conversation~\citep{moussaid_amplification_2015}.}

\newtext{Research on online content points to a third approach to this question.}
By investigating the transformations of quotations in a large corpus of US blog posts and online news stories initially collected and studied by \citet{leskovec_meme-tracking_2009}, \citet{simmons_memes_2011} and later \citet{omodei_multi-level_2012} show that even for quotations, a type of public representation that should change the least when transmitted on the Web, it is still possible to witness significant transformations.
These studies focus on the influence of the quotation source (\hbox{e.g.} news outlet \hbox{vs.} blog) or of the surrounding public space (\hbox{e.g.} quotation frequency in the corpus), and suggest diffusion-transformation models to capture the dynamics of the population of quotations.
But the cognitive features which may determine or, at least, influence these transformations, are overlooked.
\newtext{Yet cognitive and linguistic features have been used in studies not involving transformation: \citet{danescu-niculescu-mizil_you_2012}, for instance, show that particularly memorable quotations (taken from movie scripts in this case) use more distinctive words and have more common syntax than less memorable quotations; they are also the quotes that adapt best to new contexts of use.
One source of ideas to study the transformations of such quotes, then, might be the psycholinguistic literature studying word and sentence recall.}

@@

Compared to full sentences, recall of word lists provides a situation that is easier to fully explore, and that has been extensively studied.
In particular, the Deese, Roediger, and McDermott paradigm (introduced by \citealp{deese_prediction_1959}, and later popularized by \citealp{roediger_creating_1995}) has shown that it is possible to construct lists of words which reliably create the false memory of an external word related to those in the list.
This is done by using lists of words produced by free association from the target intrusion word;
the intruding recall then happens with probability nearly proportional to the average strength of semantic association between the intruding word and the words in the list.
A sizable literature studies this type of task with varying complexities in the design of the lists, a good review of which is given by \citet{zaromb_temporal_2006}.
One notable effect is that the semantic relations between words greatly influence, and correlate to, the order in which words are recalled~\citep{tulving_subjective_1962,howard_when_2002}, and that this reordering of items improves subjects' repeated recalls~\citep{tulving_subjective_1966}.
The frequency and type of intrusions in lists of random words are also influenced by associations created by the presentation of previous lists~\citep{zaromb_temporal_2006};
indeed the question of how such temporal associations (contributing to contextual information retrieval in recall) interact with the prior semantic associations of subjects (contributing to associative information retrieval) is at the center of many of these studies.

@@

Given this complexity we decided to focus on more aggregate effects, where variations of the conditions in which sentences are read and produced have a chance of being statistically smoothed out.\footnote{
Aside from our lack of control on the precise conditions of encoding and recall in our data set, the analysis techniques mentioned above are better suited to data consisting of a high number of measures over a smaller number of lists (in which case it makes sense to ask \hbox{e.g.} what proportion of intrusions come from prior lists).
As is explained further down however, our data set is shaped the opposite way: a great number of sentences, with only very few to no measures at all on each sentence.
}
If a cognitive bias in the substitution of words manifests itself with simple measures, then it will be worth applying predictive models of the substitution process in further research.

Lexical features, then, are obvious well-studied word measures that can be analyzed in aggregate.
Indeed word frequency~\citep[see][for a review]{yonelinas_nature_2002}, age-of-acquisition~\citep{zevin_age_2002}, number of phonemes~\citep[see for instance][]{rey_phoneme_1998,nickels_dissociating_2004}, and phonological neighborhood density~\citep{garlock_age--acquisition_2001} to name a few, all have known effects on word recognition or production.
More complex features based on word networks built from free association or phonological data have also been analyzed:
\citet{nelson_how_2013} for instance, show the importance of clustering coefficient in such a semantic network by studying the role it plays in a variety of recall and recognition tasks (extralist and intralist cuing, single item recognition, and primed free association).
\citet{chan_network_2010} show that pictures are named faster and with fewer mistakes when they have a lower clustering coefficient in an underlying phonological network.
\citet{griffiths_google_2007} analyze a task where subjects are asked to name the first word which comes to their mind when they are presented with a random letter from the alphabet.
The authors show that there is a link between the ease of recall of words and their authority position (pagerank) in a language-wide semantic network built from external word association data~(\citealp{austerweil_human_2012}, further develop this tool to give a parsimonious account of the fact that related words are often retrieved together from memory).

@@

On the whole, research on lexical features hints towards two antagonistic types of effects.
On one hand, part of the literature shows that recall is easier for the least "awkward" words;
those whose age of acquisition is earlier, length is smaller, semantic network position is more central --- this is particularly true in retrieval, that is in tasks where participants are asked to form spontaneous associations or utter a word in response to a given signal.
On the other hand, when the task consists in recognizing a specific item in a list, "awkward" words are actually more easily remembered, possibly as they are more informative and plausibly more discernible~\citep[see again][for a review]{yonelinas_nature_2002}.
The jury is still out as to whether reformulation alteration, that is spontaneous replacement of words when asked to rewrite a given utterance, is rather of the former or latter sort.
We also aim to shed some light on this debate, considering oddness as a dimension of the purported fitness of utterances.
